Iroko: Predictive Datacenter Congestion Control

### 2017-12-11 Documents/Report/implementation.tex
changes to design and implementation

\subsection{Topology Simulation}
We chose to use  Mininet, a python data center emulation API, to prototype Iroko. Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}. This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware. Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.

@@

\subsection{Data Collection }
Since we knew prior to testing which port connected to each host, we used different link statistics from each switch port connected to host to make predictions. This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.

@@

In addition, we used the following Linux command:
\[$tc -s qdisc show dev <interface>$ \] 
where $<interface>$ is the current host we are polling from. Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues. Each of these data provide further information to hint at the current state of the data center. Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions. Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from. Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.  These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.

@@

\subsection{Iroko}

In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem. The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ by choosing the best action $a$ to take at each time step~\cite{Sutton:1998:IRL:551283}. These values are can be tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}. Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.

\[ [0 0 0 ... 1 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.

\[bandwidth_{t+1}   <-- bandwidth_t +  a_t * bandwidth_t\]

For a reinforcement learning agent to learn a policy, a reward function is used to generate the value of a particular state, and is used to represent the goal of an agent ~\cite{Sutton:1998:IRL:551283}. We defined  a reward of 1 if the packet loss was 0 , and a reward of -1 for any loss greater than 0. M  A more intricate reward functions can be devised, and onea limitation of the current reward signal is that 0 loss can occuris also obtainable if no bandwidth is used in the current epoch which could confuse the agent.

We chose to use the deep deterministic policy gradient policy algorithm (DDPG) as described in ~\cite{DDPG} . DDPG is an actor-critic model which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.  Our agent learns using a replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}. This decision might be unnecessary for a data center where some authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and may suggest that each state can be viewed as uncorrelated inherently. We recognize these findings may also give an upper limit to the expected performance of Iroko.

To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}. Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$. Our random action is then defined as follows:

@@

>>>>

\subsection{Topology Simulation}
We chose to use  Mininet, a python data center emulation API, to prototype Iroko. Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}. This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware. Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot be deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.

@@

\subsection{Data Collection }
Since we knew prior to testing which port connected to each host, we collected different data from port connections to each end-host. This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.

@@

In addition, we used the following Linux command:
texttt{tc\ -s\ qdisc\ show\ dev\ <interface>}
where $<interface>$ is the current host we are polling from. Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues. Each of these data provide further information to hint at the current state of the data center. Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions. Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from. Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.  These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.

@@

\subsection{Reinforcement Learning Algorithm}

We chose to use the deep deterministic policy gradient policy algorithm (DDPG) almost entirely as described in ~\cite{DDPG} . DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283} which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.  Our agent learns using replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}. This decision might be unnecessary for a data center where authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and this suggests that we could assume data center data is inherently uncorrelated. We recognize these findings may also give an upper limit to the expected performance of Iroko.

To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}. Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}. Our random action is then defined as follows:

@@

\[ [0 0 0 ... 1 ... 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.

\[bandwidth_{t+1}   \leftarrow bandwidth_t +  a_t * bandwidth_t\]
